---
title: "Submission"
author: "MLG"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = TRUE, tidy = TRUE, tidy.opts = list(width.cutoff = 80))
```

## Data processing

Data is loaded below and we only select relevant columns:
\begin{itemize}
  \item \textit{EVTYPE}: Type of disaster (eg. tornado)
  \item \textit{FATALITIES}: amount of fatalities
  \item \textit{INJURIES}: amount of injuries
  \item \textit{PROPDMG}: total property damage
  \item \textit{PROPDMGEXP}: magnitude of property damage (eg. million)
  \item \textit{CROPDMG}: total crop damage
  \item \textit{CROPDMGEXP}: magnitude of crop damage
\end{itemize}

```{r, cache = TRUE}
library(dplyr)
library(ggplot2)

df <- read.csv('repdata_data_StormData.csv') %>% select(EVTYPE, FATALITIES, INJURIES, PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP)

head(df)
```

## Data analysis

### Across the United States, which types of events are most harmful with respect to population health?

Harmful events can be defined as the events that caused the most fatalities/injuries. This can be considered over a period (i.e. total sum) or on average (eg. some harmful events do not occur often but cause a lot of damage). We will use both interpretations to answer the question. $df$ is transformed below so that we retrieve all the parameters above:

```{r}
harmful_df <- df %>% mutate(EVTYPE = trimws(EVTYPE)) %>% group_by(EVTYPE) %>% summarise(count = n(), sum_injuries = sum(INJURIES), sum_fatalities = sum(FATALITIES), mean_injuries = mean(INJURIES), mean_fatalities = mean(FATALITIES)) %>% arrange(EVTYPE)
head(harmful_df)
```

\textit{Most fatalities/injuries}

```{r}
harmful_df %>% arrange(desc(sum_fatalities)) %>% top_n(10)
```
```{r}
harmful_df %>% arrange(desc(sum_injuries)) %>% top_n(10)
```
```{r}
harmful_df %>% arrange(desc(mean_fatalities)) %>% top_n(10)
```
```{r}
harmful_df %>% arrange(desc(mean_injuries)) %>% top_n(10)
```